# Methods

Overview of this is as in the README file, roughly as follows:

## Geographic area 

There are 21 cities and towns included in the Mystic River Watershed area: 

1. Burlington 
2. Lexington 
3. Belmont 
4. Watertown 
5. Arlington 
6. Winchester 
7. Woburn 
8. Reading 
9. Stoneham 
10. Medford 
11. Somerville 
12. Cambridge 
13. Boston (East Boston, Charlestown) 
14. Everett 
15. Malden 
16. Melrose 
17. Wakefield 
18. Chelsea 
19. Revere 
20. Winthrop 
21. Wilmington 


## ACS constraint tables 

ACS 5-year (2021) detailed estimates tables for Massachusetts census tracts were obtained. 

Race/Ethnicity: 

  - B04006 [People Reporting Ancestry](https://data.census.gov/table?q=B0400&g=040XX00US25&tid=ACSDT1Y2021.B04006) 
    - Collapsed by town/city: all ancestry groups with more than 1% of the population represented are retained, those with less than 1% are collapsed to 'other'
  - B03001 [Hispanic or Latino by Specific Origin](https://data.census.gov/table?q=B03001&g=040XX00US25,25$1400000) 
  
Sex and Age: 

  - B01001 [Sex by Age](https://data.census.gov/table?q=B01001&g=040XX00US25,25$1400000) 
  
```{r nice-tab, echo=FALSE, tidy=FALSE}
library(readxl)
B01001 <- read_excel("acs_variables.xlsx", sheet = "B01001")

knitr::kable(
 B01001, 
 caption = 'B01001 Variables',
 booktabs = TRUE
)
```
  
Education: 

  - B15001 [Sex by educational attainment for the population 18 years and over](https://data.census.gov/table?q=B15001&g=040XX00US25,25$1400000) 
  
```{r nice-tab, echo=FALSE, tidy=FALSE}
library(readxl)
B15001 <- read_excel("acs_variables.xlsx", sheet = "B15001")

knitr::kable(
 B15001, 
 caption = 'B15001 Variables',
 booktabs = TRUE
)
```

Householder Age and Household Income: 

  - B19307 [Age of householder by household income in the past 12 months (2021 inflation-adjusted dollars)](https://data.census.gov/table?q=B19037&g=040XX00US25,25$1400000) 
  
Householder Age and Tenure: 

  - B25007 [Tenure by Age of Householder](https://data.census.gov/table?q=B25007&g=040XX00US25,25$1400000) 
  



## PUMS data 

ACS 5-year (2017-2021) Public Use Microdata Sample (PUMS) files for *person* data and for *household* data. These data sample approximately 5% of individual responses to the ACS. Full documentation on this data is provided in the [PUMS User Guide](https://www2.census.gov/programs-surveys/acs/tech_docs/pums/2017_2021ACS_PUMS_User_Guide.pdf).

We have downloaded the PUMS data at the highest spatial resolution available: PUMA (Public Use Microdata Area), geographic units developed in the Decennial Census (here, 2010) of approximately 100,000 people per PUMA. 

```{r nice-tab, echo=FALSE, tidy=FALSE}
library(readr)
person <- read_csv("PUMA_p_variables.csv")
hhold <- read_csv("PUMA_h_variables.csv")

knitr::kable(
 person, 
 caption = 'Relevant PUMS Person Variables',
 booktabs = TRUE
)
knitr::kable(
 hhold, 
 caption = 'Relevant PUMS Household Variables',
 booktabs = TRUE
)
```

## Bring into the CO model

### Phase 1: with replacement

This we will do for basic demographics, the 13 characteristics as before

### Phase 2: without replacement

This we will do once we have the housing data Jon shared

## Exporting from the CO model

## bookdown boiler plate
Math can be added in body using usual syntax like this 

## math example

$p$ is unknown but expected to be around 1/3. Standard error will be approximated

$$
SE = \sqrt(\frac{p(1-p)}{n}) \approx \sqrt{\frac{1/3 (1 - 1/3)} {300}} = 0.027
$$

You can also use math in footnotes like this^[where we mention $p = \frac{a}{b}$].

We will approximate standard error to 0.027[^longnote]

[^longnote]: $p$ is unknown but expected to be around 1/3. Standard error will be approximated

    $$
    SE = \sqrt(\frac{p(1-p)}{n}) \approx \sqrt{\frac{1/3 (1 - 1/3)} {300}} = 0.027
    $$
