--- 
title: "Synthetic population generation for ACRES project"
author: "Flannery Black-Ingersoll and Chad Milando"
header-includes:
  - \usepackage{draftwatermark}
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [articles.bib, packages.bib, zotero.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This book describes the process and applications of using the CO model to create a synthetic population for MyWRA, and estimate the impact of climate-change adaptation efforts."
---

# Prerequisites

This book details synthetic population generation, or 'synthpop' generation, provides detailed methods and links (NOT YET PUBLISHED) to R Code to create your own synthpop using 
a combination of publicly available datasets. Please do not get this confused with Synth-Pop the British synthesizer movement, because, as Annie Lennox is quoted saying in the [NYTimes in 1984](https://timesmachine.nytimes.com/timesmachine/1984/08/03/109168.html?pageNumber=45):  

**"When you say 'synth-pop,' I immediately think of two guys with a synthesizer and a drum machine, and to me that's boring."**


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(tigris)
library(leaflet) #https://rstudio.github.io/leaflet/basemaps.html
library(sf)
library(dplyr)
library(utils)
library(readr)
# Plot tract polygons and puma polygons together
MA_pumas <- pumas(state = "MA", year = 2021)
pumas <- c("00503",
           "00505",
           "02800",
           "01000",
           "00508",
           "00507",
           "00506",
           "03302",
           "00505",
           "03306",
           "01300")
MA_pumas_mywra <- subset(MA_pumas, MA_pumas$PUMACE10 %in% pumas)
MA_tracts <- tracts(state = "MA", year = 2021)

my_dir <- "mywra_tractsbypuma/"
PUMA00503 <- readRDS(paste0(my_dir,"PUMA00503.rds"))
PUMA00503 <- unique(substr(PUMA00503$PUMACE10, 6,11))
PUMA00503_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA00503)
  
PUMA00505 <- readRDS(paste0(my_dir,"PUMA00505.rds"))
PUMA00505 <- unique(substr(PUMA00505$PUMACE10, 6,11))
PUMA00505_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA00505)

PUMA00506 <- readRDS(paste0(my_dir,"PUMA00506.rds"))
PUMA00506 <- unique(substr(PUMA00506$PUMACE10, 6,11))
PUMA00506_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA00506)

PUMA00507 <- readRDS(paste0(my_dir,"PUMA00507.rds"))
PUMA00507 <- unique(substr(PUMA00507$PUMACE10, 6,11))
PUMA00507_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA00507)

PUMA00508 <- readRDS(paste0(my_dir,"PUMA00508.rds"))
PUMA00508 <- unique(substr(PUMA00508$PUMACE10, 6,11))
PUMA00508_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA00508)

PUMA01000 <- readRDS(paste0(my_dir,"PUMA01000.rds"))
PUMA01000 <- unique(substr(PUMA01000$PUMACE10, 6,11))
PUMA01000_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA01000)

PUMA01300 <- readRDS(paste0(my_dir,"PUMA01300.rds"))
PUMA01300 <- unique(substr(PUMA01300$PUMACE10, 6,11))
PUMA01300_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA01300)

PUMA02800 <- readRDS(paste0(my_dir,"PUMA02800.rds"))
PUMA02800 <- unique(substr(PUMA02800$PUMACE10, 6,11))
PUMA02800_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA02800)

PUMA03302 <- readRDS(paste0(my_dir,"PUMA03302.rds"))
PUMA03302 <- unique(substr(PUMA03302$PUMACE10, 6,11))
PUMA03302_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA03302)

PUMA03306 <- readRDS(paste0(my_dir,"PUMA03306.rds"))
PUMA03306 <- unique(substr(PUMA03306$PUMACE10, 6,11))
PUMA03306_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA03306)


map_2 <- leaflet() %>%
  setView(lat = 42.48, lng = -71.09, zoom = 10) %>%
  addProviderTiles(
    providers$Esri.WorldTopoMap,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Topographic") %>%
   addProviderTiles(
    providers$Esri.WorldGrayCanvas,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Gray Canvas") %>%
  addProviderTiles(
    providers$Esri.WorldImagery,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Imagery") %>%
    addPolygons(
    data = PUMA00503_tracts,
    # set the color of outlines
    stroke = T,
    color = "red",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 00503 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA00505_tracts,
    # set the color of outlines
    stroke = T,
    color = "blue",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 00505 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA00506_tracts,
    # set the color of outlines
    stroke = T,
    color = "orange",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 00506 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA00507_tracts,
    # set the color of outlines
    stroke = T,
    color = "darkgreen",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 00507 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA00508_tracts,
    # set the color of outlines
    stroke = T,
    color = "magenta",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 00508 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA01000_tracts,
    # set the color of outlines
    stroke = T,
    color = "purple",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 01000 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA01300_tracts,
    # set the color of outlines
    stroke = T,
    color = "goldenrod",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 01300 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA02800_tracts,
    # set the color of outlines
    stroke = T,
    color = "tan",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 02800 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA03302_tracts,
    # set the color of outlines
    stroke = T,
    color = "pink",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 03302 Census Tracts",
    popup = ~TRACTCE
   ) %>%
  addPolygons(
    data = PUMA03306_tracts,
    # set the color of outlines
    stroke = T,
    color = "maroon",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "PUMA 03306 Census Tracts",
    popup = ~TRACTCE
   ) %>%
   addPolygons(
    data = MA_pumas_mywra,
    # set the color of outlines
    stroke = T,
    color = "black",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "All Public Use Microdata Areas",
    popup = ~NAMELSAD10
   )%>%
  # Layers control
  addLayersControl(
    baseGroups = c("ESRI World Topographic",
                   "ESRI World Gray Canvas",
                   "ESRI World Imagery"),
    overlayGroups = c("All Public Use Microdata Areas",
      "PUMA 00503 Census Tracts",
           "PUMA 00505 Census Tracts",
           "PUMA 02800 Census Tracts",
           "PUMA 01000 Census Tracts",
           "PUMA 00508 Census Tracts",
           "PUMA 00507 Census Tracts",
           "PUMA 00506 Census Tracts",
           "PUMA 03302 Census Tracts",
           "PUMA 00505 Census Tracts",
           "PUMA 03306 Census Tracts",
           "PUMA 01300 Census Tracts"),
    options = layersControlOptions(collapsed = FALSE)
  )

# side note... look how beautiful this is: -->
# leaflet() %>%
# setView(lat = 42.40, lng = -71.09713377834885, zoom = 10) %>%
# addProviderTiles(providers$Stamen.Watercolor)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
map_2
```


**What does a synthetic population mean? Well, the following demonstrates what a synthetic population would look like for a sample of one Public Use Microdata Area (PUMA # 00503).** 
N.B. This map is created using a random sampling of 25% of each respective census tract's synthetic population generated in order to make it easier to view. The number of dots therefore represents the population density in that tract. Also note that we of course do not assume people are swimming in the water bodies. Flannery needs to figure out how to generate random points without placing them in the bodies of water.

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(tigris)
library(leaflet) #https://rstudio.github.io/leaflet/basemaps.html
library(sf)
library(dplyr)
library(utils)
library(readr)
library(maptools)
library(sp)
# Plot tract polygons and puma polygons together
MA_pumas <- pumas(state = "MA", year = 2021)
pumas <- c("00503")
MA_pumas_mywra <- subset(MA_pumas, MA_pumas$PUMACE10 %in% pumas)
MA_tracts <- tracts(state = "MA", year = 2021)

my_dir <- "mywra_tractsbypuma/"
PUMA00503 <- readRDS(paste0(my_dir,"PUMA00503.rds"))
PUMA00503 <- unique(substr(PUMA00503$PUMACE10, 6,11))
PUMA00503_tracts <- subset(MA_tracts,MA_tracts$TRACTCE %in% PUMA00503)

synthpop_00503 <- readRDS("syntheticpop/PUMA00503_syntheticpop.RDS")
synthpop_00503$TRACTCE <- substr(synthpop_00503$tract, 6, 11)

# Prepare Hispanic Category for mapping:
unique(synthpop_00503$HISP)
# Recode: if 3, 7, or 1, then separate. All else are '4' for other hispanic
  # '3'=1, # .Puerto Rican "Puerto.Rican" 
  # '7'=2, # .Guatemalan "Guatemalan"  
  # '1'=3 #.Not Spanish/Hispanic/Latino (Other) "Other"
synthpop_00503$HISP_FACTOR <- ifelse(synthpop_00503$HISP == 3, "Puerto Rican", "Other Spanish/Hispanic/Latino")
synthpop_00503$HISP_FACTOR <- ifelse(synthpop_00503$HISP == 7, "Guatemalan", synthpop_00503$HISP_FACTOR)
synthpop_00503$HISP_FACTOR <- ifelse(synthpop_00503$HISP == 1, "Not Spanish/Hispanic/Latino", synthpop_00503$HISP_FACTOR)
table(synthpop_00503$HISP_FACTOR)
# For loop to iterate through TRACTCE
this_PUMA_all_TRACTCE <- unique(PUMA00503_tracts$TRACTCE)
all_points_with_data <- vector(mode = "list", length = length(this_PUMA_all_TRACTCE))
names(all_points_with_data) <- this_PUMA_all_TRACTCE
sampled_points_with_data <- vector(mode = "list", length = length(this_PUMA_all_TRACTCE))
names(sampled_points_with_data) <- this_PUMA_all_TRACTCE

for (this_tract in this_PUMA_all_TRACTCE){
  subset_tract <- subset(PUMA00503_tracts, TRACTCE == this_tract)
  subset_synthpop <- subset(synthpop_00503, TRACTCE == this_tract)
  num_points <- nrow(subset_synthpop)
  if(num_points == 0) next
  # Sample num_points random points within the polygon
  points <- sf::st_sample(subset_tract, size=num_points)
  # Convert from sfc to sf
  points_sf <- points %>%
               st_sf %>%
               st_cast
  points_sf$ID <- 1:num_points
  subset_synthpop$ID <- points_sf$ID
  points_with_data <- sp::merge(points_sf, subset_synthpop, by = "ID")
  # Generate a random sample of 25% of the data
  set.seed(1)
  sample_points_with_data <- points_with_data %>%
                             slice_sample(prop = 0.25)
  all_points_with_data[[this_tract]] <- points_with_data
  sampled_points_with_data[[this_tract]] <- sample_points_with_data
}

# Set your own colors manually:

pal <- colorFactor(
  palette = c( "gold1",
  "skyblue2", "#FB9A99",
  "palegreen2"),
  domain = synthpop_00503$HISP_FACTOR
)
# Create the underlying map
map3 <- leaflet() %>%
    setView(lat = 42.46566388574347, lng = -71.28568709503587, zoom = 11) %>%
    addProviderTiles(
    providers$Esri.WorldTopoMap,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Topographic") %>%
   addProviderTiles(
    providers$Esri.WorldGrayCanvas,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Gray Canvas") %>%
  addProviderTiles(
    providers$Esri.WorldImagery,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Imagery")%>%
  addPolygons(
    data = MA_pumas_mywra,
    # set the color of outlines
    stroke = T,
    color = "black",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "All Public Use Microdata Areas",
    popup = ~NAMELSAD10
   )%>%
  addPolygons(
    data = PUMA00503_tracts,
    # set the color of outlines
    stroke = T,
    color = "black",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = .8,
    group = "All PUMA00503 Tracts",
    popup = ~TRACTCE
   )
# Loop over various polygons to include.
overlay_vector <- vector(mode = "list")
subset_sample <- subset(sampled_points_with_data, !names(sampled_points_with_data) == "359302")
for (tract_number in 1:length(subset_sample)) {
    this_tract_data <- subset_sample[[tract_number]]
    overlay_vector[[tract_number]] <- paste0("Tract #", (PUMA00503[tract_number])) 
    map3 <- map3 %>%
            addCircles(
            data = this_tract_data,
            # set the color of outlines
            stroke = F,
            color = ~pal(this_tract_data$HISP_FACTOR), 
            fillOpacity = .8,
            fill = T,
            # set the opacity of the outline
            # set the stroke width in pixels
            weight = .6,
            radius = .1,
            group = paste0("Tract #", (PUMA00503[tract_number]))
            )
}

map3 <- map3 %>%
  addLayersControl(
    baseGroups = c("ESRI World Topographic",
                   "ESRI World Gray Canvas",
                   "ESRI World Imagery"),
    overlayGroups = c("All Public Use Microdata Areas",
                      "All PUMA00503 Tracts",
                      overlay_vector),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend("topleft",
  pal = pal,
  values = synthpop_00503$HISP_FACTOR,
  title = "Hispanic Ancestry Category",
  opacity = .8)

# ACS Chloropleth
ACS_data <- read_csv("ACS_10_5YR_B03001_mywra_PUMA00503.csv")
ACS_data <- subset(ACS_data, ACS_data$Total != 0)
ACS_data$TRACTCE <- substr(ACS_data$Id, 15, nchar(ACS_data$Id))
ACS_data$Guat_perc <- ifelse(ACS_data$Guatemalan != 0, 100*(ACS_data$Guatemalan/ACS_data$Total), 0)
ACS_data_geom <- sp::merge(PUMA00503_tracts, ACS_data, by = "TRACTCE")

# Color palette
library(RColorBrewer)
mybins <- c(0,5,10,15,20,25,30,35)
pal2 <- colorBin( palette="YlOrRd",
                  alpha = T, 
                  domain=ACS_data_geom$Guat_perc, 
                  #na.color="transparent", 
                  bins=mybins)
# TODO why is this so transparent?!
# Basic choropleth with leaflet?
map4 <- leaflet() %>% 
   setView(lat = 42.46566388574347, lng = -71.28568709503587, zoom = 11) %>%
    addProviderTiles(
    providers$Esri.WorldTopoMap,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Topographic") %>%
   addProviderTiles(
    providers$Esri.WorldGrayCanvas,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Gray Canvas") %>%
  addProviderTiles(
    providers$Esri.WorldImagery,
    options = providerTileOptions(opacity = 0.75),
    group = "ESRI World Imagery")%>%
  addPolygons(
    data = MA_pumas_mywra,
    # set the color of outlines
    stroke = T,
    color = "black",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = 1.4,
    group = "All Public Use Microdata Areas",
    popup = ~NAMELSAD10
   )%>%
  addPolygons(
    data = PUMA00503_tracts,
    # set the color of outlines
    stroke = T,
    color = "black",
    fill = F,
    # set the opacity of the outline
    # set the stroke width in pixels
    weight = .8,
    group = "All PUMA00503 Tracts",
    popup = ~TRACTCE
   )%>%
  addPolygons(
    data = ACS_data_geom,
    fillColor = ~pal2(Guat_perc),
    opacity = .95,
    stroke=FALSE,
    group = "Guatemalan Population %")%>%
  addLayersControl(
    baseGroups = c("ESRI World Topographic",
                   "ESRI World Gray Canvas",
                   "ESRI World Imagery"),
    overlayGroups = c("All Public Use Microdata Areas",
                      "All PUMA00503 Tracts",
                      "Guatemalan Population %"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
  "topleft",
  pal = pal2,
  values = ACS_data_geom$Guat_perc,
  title = "Guatemalan Population %",
  opacity = .8)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
map3
```

**We can compare this population to the census tract level statistics provided by ACS. Let's take a look at the Guatemalan population by tract, for example. We see a similar concentration in the southern tracts of the PUMA.**

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
map4
# 
# # Redlining
# library(rgdal)
# redlining <- readOGR( 
#   dsn= "redlining_shapefile/redlining_shp/" , 
#   layer="MA_Redlining",
#   verbose=FALSE
# )
# redlining_sf <- redlining %>%
#                st_as_sf %>%
#                st_cast
# unique(redlining_sf$holc_grade)
# subset_redline <- subset(redlining_sf, redlining_sf$layer %in% c("Waltham",
# "Concord",
# "Lexington",
# "Lincoln",
# "Bedford",
# "Burlington",
# "Watertown",
# "Weston"))
# # Palette
# subset_redline$holc_grade_num <- ifelse(subset_redline$holc_grade == "A",1,4)
# subset_redline$holc_grade_num <- ifelse(subset_redline$holc_grade == "B",2,subset_redline$holc_grade_num)
# subset_redline$holc_grade_num <- ifelse(subset_redline$holc_grade == "C",3,subset_redline$holc_grade_num)
# pal3 <- colorFactor(
#   palette = c( "pink",
#   "red2", "red3",
#   "red4"),
#   domain = subset_redline$holc_grade_num
# )
# 
# 
# # Map 5
# map5 <- leaflet() %>% 
#    setView(lat = 42.46566388574347, lng = -71.28568709503587, zoom = 11) %>%
#     addProviderTiles(
#     providers$Esri.WorldTopoMap,
#     options = providerTileOptions(opacity = 0.75),
#     group = "ESRI World Topographic") %>%
#    addProviderTiles(
#     providers$Esri.WorldGrayCanvas,
#     options = providerTileOptions(opacity = 0.75),
#     group = "ESRI World Gray Canvas") %>%
#   addProviderTiles(
#     providers$Esri.WorldImagery,
#     options = providerTileOptions(opacity = 0.75),
#     group = "ESRI World Imagery")%>%
#   addPolygons(
#     data = MA_pumas_mywra,
#     # set the color of outlines
#     stroke = T,
#     color = "black",
#     fill = F,
#     # set the opacity of the outline
#     # set the stroke width in pixels
#     weight = 1.4,
#     group = "All Public Use Microdata Areas",
#     popup = ~NAMELSAD10
#    )%>%
#   addPolygons(
#     data = PUMA00503_tracts,
#     # set the color of outlines
#     stroke = T,
#     color = "black",
#     fill = F,
#     # set the opacity of the outline
#     # set the stroke width in pixels
#     weight = .8,
#     group = "All PUMA00503 Tracts",
#     popup = ~TRACTCE
#    )%>%
#   addPolygons(
#     data = subset_redline,
#     stroke = T,
#     color = "black",
#     fill = T
#     # fillColor = ~pal3(holc_grade_num),
#     # opacity = 1,
#     # stroke=FALSE
#     )%>%
#   addLayersControl(
#     baseGroups = c("ESRI World Topographic",
#                    "ESRI World Gray Canvas",
#                    "ESRI World Imagery"),
#     overlayGroups = c("All Public Use Microdata Areas",
#                       "All PUMA00503 Tracts",
#                       "HOLC Grade"),
#     options = layersControlOptions(collapsed = FALSE)
#   ) %>%
#   addLegend(
#   "bottomleft",
#   pal = pal3,
#   values = ~subset_redline$holc_grade_num,
#   title = "HOLC Grade",
#   opacity = .8)
# 
# map5
```

